% report.tex
% LaTeX report template
\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}

\title{Your Report Title}
\author{Your Name}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
Your abstract goes here.
\end{abstract}

\section{Introduction}
\subsection{Whittaker Smoothing vs. Asymmetric Least Squares (AsLS) for Spectroscopy}

\textbf{Whittaker smoothing} is a classic method for smoothing data, formulated as a minimization problem:
\begin{equation*}
	\min_z \sum_i (y_i - z_i)^2 + \lambda \sum_i (D^2 z_i)^2
\end{equation*}
where $y$ is the observed signal, $z$ is the smoothed signal, $D^2$ is the second difference operator, and $\lambda$ controls the trade-off between fidelity to the data and smoothness. Whittaker smoothing is linear, treats positive and negative residuals equally (symmetric), and has a single trade-off parameter $\lambda$. The solution is closed-form, and the bias--variance trade-off is easily visualized using the L-curve.

\textbf{Asymmetric Least Squares (AsLS)} modifies only the data fidelity term by introducing weights:
\begin{equation*}
	\min_z \sum_i w_i (y_i - z_i)^2 + \lambda \sum_i (D^2 z_i)^2
\end{equation*}
where $w_i = p$ if $y_i > z_i$ and $w_i = 1-p$ if $y_i < z_i$, with $p$ typically close to zero. The weights depend on the current solution $z$, making the problem iterative and asymmetric. AsLS actively ignores peaks by down-weighting positive residuals, focusing on fitting the lower envelope of the signal---ideal for baseline removal in spectroscopy.

\textbf{Key differences:}
\begin{itemize}
	\item \textit{Residual weighting:} Whittaker is symmetric; AsLS is asymmetric.
	\item \textit{Solution:} Whittaker is solved in one shot; AsLS is iterative.
	\item \textit{Interpretation:} Whittaker fits all data but penalizes wiggles; AsLS fits only the baseline, ignoring peaks.
\end{itemize}

\textbf{When to use which:}
\begin{itemize}
	\item \textit{Noise reduction or trend extraction:} Whittaker
	\item \textit{Baseline removal in spectroscopy:} AsLS
\end{itemize}

For AsLS, $\lambda$ is typically chosen by heuristics (e.g., $\lambda \sim 10^4$--$10^8$ for spectroscopy) and visual inspection.In summary, AsLS is not simply Whittaker smoothing; it uses Whittaker as a subproblem but adds asymmetric, adaptive weighting, making it more suitable for baseline correction in spectroscopy.
Write your introduction here.



\section{Methods}
Describe your methods here.

\section{Results}
Present your results here.

\section{Discussion}
Discuss your findings here.

\section{Conclusion}
Summarize your conclusions here.

\bibliographystyle{plain}
\bibliography{references}

\end{document}
